Toward a novel piezoelectric energy harvester for earcanal dynamic motion exploitation, using a bistable resonator cycled by coupled hydraulic valves made of collapsed flexible tubes

### 2022-10-31 article_system.tex
Après corrections Jérémie

% for cutting words
\usepackage[utf8]{inputenc}
% a font close to that chosen by Elsevier
\usepackage[bitstream-charter]{mathdesign}
\usepackage{mathtools, cuted}
% for accentuated characters
%\usepackage[T1]{fontenc}
\geometry{top=2cm,left=1.5cm,bottom=2cm,right=1.5cm}
\usepackage{multicol}
% For treating ps:
\usepackage{psfrag}
% For subfigures
\usepackage{subcaption}
\usepackage{placeins}
% For tables
\usepackage{array}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{colortbl}
\usepackage{tabulary}
\usepackage{etoolbox}
% For citations
\usepackage[numbers]{natbib}
%\biboptions{longnamesfirst,angle,semicolon}
%\citetstyle{nature}
% For color
\usepackage{color}
\usepackage{xcolor}
% For spacing
\usepackage{setspace}
%\doublespacing
\singlespacing
%\onehalfspacing
% For maths
\usepackage{amsmath, amsfonts}
\usepackage[squaren,ashgrey]{SIunits}
% For Line numbering
\usepackage[pagewise,modulo]{lineno}
%\linenumbers
% For special characters
\usepackage{pifont}
%
\usepackage[pagebackref,colorlinks=True]{hyperref}
\usepackage{wasysym}
\usepackage{color}
\usepackage{cleveref}
\crefformat{figure}{(Fig.~#2#1#3)}
\usepackage{hyperref}
\usepackage[hyperpageref]{backref}
\usepackage{lipsum}
\usepackage{booktabs} % To thicken table lines
% \usepackage{widetext}
\usepackage{float}
\usepackage{empheq}
\usepackage{textcomp}
\usepackage{multicol}
\usepackage{graphicx}

%----------------------------------------------------------------------
%                     Commands
%----------------------------------------------------------------------
\renewcommand*{\overrightarrow}[1]{\vbox{\halign{##\cr \tiny\rightarrowfill\cr\noalign{\nointerlineskip\vskip1pt}$#1\mskip2mu$\cr}}}
%% Degree character
\DeclareTextSymbol{\degr}{T1}{6}
\newcommand{\degre}[0]{\degr C}
\DeclareUnicodeCharacter{2212}{-}
% MPam^0.5 character
\newcommand{\mpa}[0]{\mbox{MPa}\sqrt{\mbox{m}}}
%mettre de nouilles sous les tenseurs
\newcommand{\ve}[1]{{\underline{#1}}}
\def\ten#1{\oalign{$#1$\crcr\hidewidth$\scriptscriptstyle\sim$\hidewidth}} 
% derivee seconde
\newcommand{\fda}[2]{\displaystyle \frac{\partial\,{#1}}{\partial\, {#2}}}
% et al. for citations
\newcommand{\etal}[0]{{\em et al.}}
\renewcommand*\contentsname{\large{Table of contents}}
\def \hfillx {\hspace*{ -\linewidth} \hfill} % remplir espacement horizontal entre deux sous-figures de façon à remplir toute la largeur du texte.
% \renewcommand{\degree}[1]{${#1}^\circ$} % °

%----------------------------------------------------------------------
%                     Colors
%----------------------------------------------------------------------
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
%----------------------------------------------------------------------
%                     Pre-Document
%----------------------------------------------------------------------
\journal{Applied Energy}
%----------------------------------------------------------------------
%                     Document
%----------------------------------------------------------------------
\begin{document}

@@

\title{Ear canal dynamic motion piezoelectric energy harvester using a bistable resonator cycled by coupled hydraulic switches made of collapsed flexible tubes.}

@@

\address[symme]{Laboratoire SYMME - Université Savoie Mont Blanc, 7 Chemin de Bellevue, 74940, Annecy}
\address[critias]{Université du Québec - École de technologie supérieure, 1100 Rue Notre-Dame Ouest, Montréal, QC, H3C 1K3}

@@

\author[critias]{Michel DEMUYNCK}
\author[critias]{Aidin DELNAVAZ}
\author[critias]{Jérémie VOIX}

\begin{abstract}
%% Text of abstract
\end{abstract}

\begin{keyword}
%% Keywords
\end{keyword}

@@

		BB                       & Buckled beam                                           \\

@@

		GB                       & Guide beam                                             \\

@@

The growing use of wireless devices and the miniaturization of electronic circuits have led to significant progress on the energy consumption of mobile devices on the human body environment. Energy harvesting methods have been studied in order to complement their power supply and enhance the autonomy of the batteries. In-ear devices such as hearing aids and cochlear implants are powered by disposable cells or rechargeable batteries. The energy consumption of the best integrated devices in the literature rises up to $17$J for a 10 hours daily use \cite{Scherer2019,Yip2015,Kulah2022}. Woodruff \emph{et al.} underlined that patients using hearing aids sometimes struggle to change and select batteries for their devices \cite{Woodruff2021}. Another statistical study revealed that the majority of hearing aids and cochlear implants users prefer disposable batteries for the long autonomy, but would have liked rechargeable solutions if the battery cycle was sufficient to complete the day \cite{PracticesAudiology2016}. Also, the long term use makes the rechargeable solutions a more economical and more ecological choice compared to the disposable cells. These arguments motivated the development of energy harvesting systems.

The common energy sources are solar or wind energy but their availability in the low size environment of the application context makes them not suitable for exploitation. The activities of the human body generate a considerable amount of energy of different natures. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, the kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, the body heat thermal energy \cite{Kim2014}, the strain energy from the skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the ear canal mechanical deformation as it is directly located in the area of the application. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements changes de ear canal geometry during mastication. Carioli \emph{et al.} then used a customized earplug molding technic with a 3D scanner to reveal a global bending movement and a local compression area in the ear canal during mastication \cite{Carioli2016}. These two mechanical deformations represent a potential energy source that can be exploited.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug has been molded on the subject in order to maximize the contact surface with the ear canal. The harvester is capable of exploiting both the flexion and the compression energy sources. The experimental prototype was able to extract $44$µJ from a mastication cycle. The soft materials facilitate the bio-integration of the harvester but the PVDF low electromechanical coupling coefficient could have led to poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy is extracted by a liquid filled earplug whose internal volume vary as a result of the radial compression of the ear canal internal walls during the jaw movements. This variation put into motion a moving magnet translating in a water column surrounded by a coil. The generated energy experimentally amounts to 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57Hz) and a small scale application \cite{Kulah2008,Priya2017}. However, the hydraulic transmission allows the use of more complex and voluminous technological solutions since the energy becomes available outside the small ear canal volume.

Based on the analysis of the literature, we propose a new solution optimizing the energy conversion efficiency to maximize the harvested energy from the ear canal dynamic motion. The next section states on the hydraulic energy source characterization and on the harvesting strategy that we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING} details the global system modeling and section \ref{sec:NUMERICAL MODEL AND SIMULATIONS} presents the numerical simulation of the theoretical behavior of the harvester. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS OF THE ELECTROMECHANICAL CONVERTER} then exposes the experimental validation performed on the electromechanical converter stage and section \ref{sec:EXPERIMENTAL APPROACH FOR THE HV DESIGN} shows the experimental approach for the design of the hydraulic valves for the system cycling. Then, section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model and discusses this work major contributions and the harvester potential. Finally, a conclusion is stated.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{ENERGY SOURCE AND HARVESTING STRATEGY}
\label{sec:THE ENERGY SOURCE AND HARVESTING STRATEGY}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The energy source}	
	\label{The energy source}
    %/////////////////////////////////////////////
Human tissue requires soft flexible materials in order to extract the ear canal deflection energy without causing discomfort. However, flexible transducers admit low energy conversion efficiency as a part of the extracted energy is dissipated in the elastic non-active material. Additionally, the low frequency of mastication makes it difficult to use resonant structures that generally admit better conversion efficiencies than low frequency solutions \cite{Ashraf2011}. Thus, the traditional transducing methods are not suitable enough for the application.

To date, the best solution proposed in the literature consists on using soft materials as the piezoelectric PVDF earplug previously introduced \cite{Delnavaz2013}. The low mechanical impedance of the soft tissue in fact limits the use of more effective rigid piezoelectric materials.

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but allows to transmit the energy outside the ear canal and benefit so from larger effective space. The latter is in fact essential as it becomes then possible to implement an impedance adaption stage and more effective transducing methods that could not fit directly inside the ear canal restricted volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is as much better as the effective volume of transducing material is high \cite{Marin2011}. The present work is then based on the exploitation of the ear canal by a liquid filled earplug. As energy source, it can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the ear canal wall. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation, with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Ear canal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behavior has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Then, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ et $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12kPa$ and $(\Delta V_{ear})_{max}=60$mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.9$mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$mJ available per day, per ear. This amount of energy could theoretically enhance supply $23.5$\% of the $17$J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot (\Delta V_{ear})_{max} = 0.9\text{mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{Harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. The power density of electromagnetic transducers decreases as much as the effective volume of transducing material and the frequency of the energy source decrease \cite{Priya2017,Kulah2008}. Also, the fabrication and the integration of the coil and the moving magnet can come out to be difficult as the gap between the two components has to be carefully managed to prevent non-linear unpredicted responses \cite{Caruntu2001}.

Piezoelectric ceramics appear like good candidates for mechanical energy transduction at small scales, but their high stiffness can limit their use on soft materials as human tissues. Mechanical amplifiers like flextensional structures can nevertheless be used to adapt the stress/displacement levels between the energy source and the transducing stiff material \cite{Abdelnaby2016}. Also, the best energy conversion efficiency for piezoelectric ceramics is achieved for at the transducer resonance. As the mean mastication frequency is estimated to \mbox{$1.57$Hz}, the harvesting system could benefit from a frequency-up conversion stage \cite{Ashraf2011,Peng2021} to increase the frequency of the energy source and so the energy conversion efficiency of the transducer. A majority of the frequency-up systems with piezoelectric transducers use mechanical stops with piezoelectric ceramic cantilevers to reach high frequencies with a good mechanical coupling \cite{Edwards2013,Gu2011,Lee2007}. Stoppers however dissipate energy. Alternative solutions use mechanically bistable resonators (BR) \cite{Vocca2012}. Recent works led to the development and optimization of bistable structures integrating stacked piezoelectric ceramics inside flextensional elastic structures \cite{Huguet2017}. They show enhanced performances due to high electromechanical coupling. The electromechanical converter, composed of a BR with piezoelectric ceramics in a flextensional structure, theoretically offers higher energy conversion efficiency compared to low frequency solutions and/or flexible electromechanical transducers \cite{Abdelnaby2016,Peng2021}.

The hydraulic-mechanic interface can be a hydraulic cylinder (HC). It can store the energy in the frequency-up converter by actuating it at the frequency of human mastication. The amount of extractable energy from the hydraulic source will then depend on the mechanical stage behavior. Before being converted in electricity, the energy is in fact stored in a mechanical potential reservoir. During the low frequency actuation, the forces applied on each side of the HC piston are equal. Equation \ref{eq:hydraulic-mechanic_interface} models the hydraulic-mechanic interface with $x_{m}$ and $ S_{hc}$ being respectively the dynamic mass position on the $\vec{x}$ axis and the HC internal section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={19.7cm 0cm 0cm 13cm},clip, width=0.7\linewidth]{figures/hydraulic_mechanic_interface.pdf}
	\caption{Hydraulic source exploitation depending on the mechanical load}
	\label{fig:hydraulic_mechanic_interface}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Therefore, the potential energy stored in a linear resonator from the ideal hydraulic source previously described can be expressed as follows :
\begin{equation}
	Ep_{lin} = \dfrac{K_s\ x_{m}^2}{2} = \dfrac{\Delta p_{ear} \Delta V_{ear}}{2}
	\label{eq:best_linear_soution}
\end{equation}
Thus, the linear solution can theoretically exploit/store a maximum of $50\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$).

>>>>

% for cutting words
\usepackage[utf8]{inputenc}
% a font close to that chosen by Elsevier
\usepackage[bitstream-charter]{mathdesign}
\usepackage{mathtools, cuted}
% for accentuated characters
%\usepackage[T1]{fontenc}
\geometry{top=2cm,left=1.5cm,bottom=2cm,right=1.5cm}
\usepackage{multicol}
% For treating ps:
\usepackage{psfrag}
% For subfigures

\usepackage{subcaption}
\usepackage{placeins}
% For tables
\usepackage{array}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{colortbl}
\usepackage{tabulary}
\usepackage{etoolbox}
% For citations
\usepackage[numbers]{natbib}
%\biboptions{longnamesfirst,angle,semicolon}
%\citetstyle{nature}
% For color
\usepackage{color}
\usepackage{xcolor}
% For spacing
\usepackage{setspace}
%\doublespacing
\singlespacing
%\onehalfspacing
% For maths
\usepackage{amsmath, amsfonts}
\usepackage[squaren,ashgrey]{SIunits}
% For Line numbering
\usepackage[pagewise,modulo]{lineno}
%\linenumbers
% For special characters
\usepackage{pifont}
%
\usepackage[pagebackref,colorlinks=True]{hyperref}
\usepackage{wasysym}
\usepackage{color}
\usepackage{cleveref}
\crefformat{figure}{(Fig.~#2#1#3)}
\usepackage{hyperref}
\usepackage[hyperpageref]{backref}
\usepackage{lipsum}
\usepackage{booktabs} % To thicken table lines
% \usepackage{widetext}
\usepackage{float}
\usepackage{empheq}
\usepackage{graphicx}

%----------------------------------------------------------------------
%                     Commands
%----------------------------------------------------------------------
\renewcommand*{\overrightarrow}[1]{\vbox{\halign{##\cr \tiny\rightarrowfill\cr\noalign{\nointerlineskip\vskip1pt}$#1\mskip2mu$\cr}}}
%% Degree character
\DeclareTextSymbol{\degr}{T1}{6}
\newcommand{\degre}[0]{\degr C}
\DeclareUnicodeCharacter{2212}{-}
% MPam^0.5 character
\newcommand{\mpa}[0]{\mbox{MPa}\sqrt{\mbox{m}}}
%mettre de nouilles sous les tenseurs
\newcommand{\ve}[1]{{\underline{#1}}}
\def\ten#1{\oalign{$#1$\crcr\hidewidth$\scriptscriptstyle\sim$\hidewidth}} 
% derivee seconde
\newcommand{\fda}[2]{\displaystyle \frac{\partial\,{#1}}{\partial\, {#2}}}
% et al. for citations
\newcommand{\etal}[0]{{\em et al.}}
\renewcommand*\contentsname{\large{Table of contents}}
\def \hfillx {\hspace*{ -\linewidth} \hfill} % remplir espacement horizontal entre deux sous-figures de façon à remplir toute la largeur du texte.
%----------------------------------------------------------------------
%                     Colors
%----------------------------------------------------------------------
\definecolor{ashgrey}{rgb}{0.7, 0.75, 0.71}
%----------------------------------------------------------------------
%                     Pre-Document
%----------------------------------------------------------------------
\journal{Applied Energy}
%----------------------------------------------------------------------
%                     Document
%----------------------------------------------------------------------
\begin{document}

@@

\title{Toward a novel piezoelectric energy harvester for earcanal dynamic motion exploitation, using a bistable resonator cycled by coupled hydraulic valves made of collapsed flexible tubes}

@@

\address[symme]{Laboratoire SYMME - Université Savoie Mont Blanc, 7 Chemin de Bellevue, 74940, Annecy}
\address[critias]{Université du Québec - École de technologie supérieure, Montréal, QC, H3C 1K3}

@@

%\author[critias]{Michel DEMUYNCK}
\author[critias]{Aidin DELNAVAZ}
\author[critias]{Jérémie VOIX}

\begin{abstract}
Scavenging energy from the earcanal dynamic motion during jaw motion may be a practical way to enhance the battery autonomy of hearing aids. One of the main challenges is to optimize the extracted energy while coping with the constraints associated with soft human tissues and restricted earcanal volume. We propose a new concept of harvester made of a liquid filled earplug transferring the energy outside the earcanal to a generator. It is composed oa hydraulic amplifier, two hydraulic cylinders actuating a bistable resonator raising the source frequency driving an amplified piezoelectric transducer to generate electricity. The cycling of the resonator is achieved using two innovative flexible hydraulic valves based on the buckling of flexible tubes.\\
A multiphysics coupled model is established in order to determine the system operation requirements and to evaluate its theoretical performance. It exhibits a theoretical energy conversion efficiency of 85\%. The electromechanical performance of the resonator couple to the piezoelectric transducer as well as the hydraulic behaviour of the valves are experimentally investigated. The global model is updated with the experimental data to ameliorate its predictability for a further optimized design. The energy losses are identified, which allow an entire design proposed to improve the experimental energy conversion efficiency to 26\%.
\end{abstract}

\begin{keyword}
Energy harvesting, frequency-up, earcanal, multiphysics, hydraulic valve
\end{keyword}

@@

		BB                       & Buckled blade                                           \\

@@

		GB                       & Guide blade                                             \\

@@

The growing use of wireless devices and the miniaturization of electronic circuits have led to significant progress on the energy consumption of mobile devices on the human body environment. Energy harvesting methods have been studied in order to complement their power supply and enhance the autonomy of the batteries. In-ear devices such as hearing aids and cochlear implant are powered by disposable cells or rechargeable batteries. The energy consumption of the best integrated devices in the literature rises up to $17$J for a 10 hours daily use \cite{Scherer2019,Yip2015,Kulah2022}. Woodruff \emph{et al.} underlined that patients using hearing aids sometimes struggle to change and select batteries for their devices \cite{Woodruff2021}. Another statistical study revealed that the majority of hearing aids and cochlear implants users prefer disposable batteries for the long autonomy, but would have liked rechargeable solutions if the battery cycle was sufficient to complete the day \cite{PracticesAudiology2016}. Also, the long-term use makes the rechargeable solutions a more economical and more ecological choice compared to the disposable cells. These arguments motivated the development of energy harvesting systems.

The common energy sources are solar or wind energy but their availability in the low-size environment of the application context makes them not suitable for exploitation. The activities of the human body generate a considerable amount of energy of different natures. Among them, the electrochemical energy from the inner ear \cite{Mercier2012}, the kinetic energy from walking or head movements \cite{Azimi2021,Smilek2016}, the body heat thermal energy \cite{Kim2014}, the strain energy from the skin deformation \cite{Jin2021} could be harvested for in-ear applications. An interesting energy source for cochlear implants or hearing aids appears to be the earcanal mechanical deformation as it is directly located in the area of the application. Delnavaz \emph{et al.} first showed in 2012 that the jaw movements changes earcanal geometry during mastication. Carioli \emph{et al.} then used a customized earplug moulding technique with a 3D scanner to reveal a global bending movement and a local compression area in the earcanal during mastication \cite{Carioli2016}. These two mechanical deformations represent a potential energy source that can be exploited.

Delnavaz \emph{et al.} developed a flexible piezoelectric harvester made of a silicon earplug with a PVDF layer \cite{Delnavaz2013}. The earplug has been moulded on the subject in order to maximize the contact surface with the earcanal. The harvester is capable of exploiting both the flexion and the compression energy sources. The experimental prototype was able to extract $44$µJ from a mastication cycle. The soft materials facilitate the bio-integration of the harvester but the PVDF low electromechanical coupling coefficient could have led to poor energy conversion efficiency. A hydro-electromagnetic energy harvester was also developed by the same researchers \cite{Delnavaz2012}. The energy is extracted by a liquid filled earplug whose internal volume varies as a result of the radial compression of the earcanal internal walls during the jaw movements. This variation put into motion a moving magnet translating in a water column surrounded by a coil. The generated energy experimentally amounts to 0.2µJ per mastication cycle. The prototype appeared to be less efficient than the PVDF earplug. This could be explained by the poor power generation of the electromagnetic transduction for a low frequency energy source (1.57\,Hz) and a small-scale application \cite{Kulah2008,Priya2017}. However, the hydraulic transmission allows the use of more complex and voluminous technological solutions since the energy becomes available outside the small earcanal volume.

Based on the analysis of the literature, we propose a new solution optimizing the energy conversion efficiency to maximize the harvested energy from the earcanal dynamic motion. The next section states on the hydraulic energy source characterization and on the harvesting strategy that we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING} details the global system modelling and section \ref{sec:NUMERICAL MODEL AND SIMULATIONS} presents the numerical simulation of the theoretical behaviour of the harvester. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS OF THE ELECTROMECHANICAL CONVERTER} then exposes the experimental validation performed on the electromechanical converter stage and section \ref{sec:EXPERIMENTAL APPROACH FOR THE HV DESIGN} shows the experimental approach for the design of the hydraulic valves for the system cycling. Then, section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model and discusses this work major contributions and the harvester potential. Finally, a conclusion is stated.

%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{ENERGY SOURCE AND HARVESTING STRATEGY}
\label{sec:THE ENERGY SOURCE AND HARVESTING STRATEGY}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The energy source}	
	\label{The energy source}
    %/////////////////////////////////////////////
Human tissue requires soft flexible materials in order to extract the earcanal deflection energy without causing discomfort. However, flexible transducers admit low energy conversion efficiency as a part of the extracted energy is dissipated in the elastic non-active material. Additionally, the low frequency of mastication makes it difficult to use resonant structures that generally admit better conversion efficiencies than low frequency solutions \cite{Ashraf2011}. Thus, the traditional transducing methods are not suitable enough for the application.

To date, the best solution proposed in the literature consists of using soft materials as the piezoelectric PVDF earplug previously introduced \cite{Delnavaz2013}. The low mechanical impedance of the soft tissue in fact limits the use of more effective rigid piezoelectric materials.

The hydraulic extraction strategy proposed in \cite{Delnavaz2012} is less effective than the piezoelectric earplug but allows to transmit the energy outside the earcanal and benefit so from larger effective space. The latter is in fact essential as it becomes then possible to implement an impedance adaptation stage and more effective transducing methods that could not fit directly inside the earcanal limited volume. In fact, Marin \emph{et al.} demonstrated that the power generated by piezoelectric or electromagnetic transducers is as much better as the effective volume of transducing material is high \cite{Marin2011}. The present work is then based on the exploitation of the earcanal by a liquid filled earplug. As an energy source, it can be assimilated to a micro pump characterized by its volume variation $V_{ear}$ and pressure variation $p_{ear}$. In order to optimize the energy extraction, the earplug must fit the earcanal wall. Turcot \emph{et al.} demonstrated that the initial pressurization must reach a minimum of $14$kPa to achieve a good fit \cite{TURCOT2011}. Bouchard-Roy \emph{et al.} studied the dynamic pressure variation inside the earplug for different initial pressurization levels \cite{Bouchard-Roy2020}. A maximum of $\text{max}(\Delta p_{ear})=12$kPa of dynamic pressure variation, with a $28$kPa of initial pressurization, has been noted during the mastication for seven subjects. Besides, the temporal evolution of the volume variation $\Delta V_{ear}$ has been estimated in \cite{Delnavaz2012} and the result for a mastication cycle is presented on figure \ref{fig:deltaV_ear}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={20.5cm 0cm 0cm 10.8cm},clip, width=0.4\textwidth]{figures/deltaV_ear.pdf}
	\caption{Earcanal volume variation for one mastication cycle}
	\label{fig:deltaV_ear}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The earplug operation characteristic behaviour has not been studied yet. Thus, the evolution of the volume variation $\Delta V_{ear}$ is not well known as a function of the pressure variation $\Delta p_{ear}$. Then, considering the pressure and volume variation levels from \cite{Delnavaz2012} and \cite{Bouchard-Roy2020}, the equation \ref{eq:max_hydraulic_energy} approximates an upper limit for the extractable hydraulic energy with the maximum values of $\Delta p_{ear}$ et $\Delta V_{ear}$ that have been recorded, i.e. $p_c=12$\,kPa and $(\Delta V_{ear})_{max}=60$\,mm$^3$. $p_c$ stands for the comfort pressure by assuming that the user could feel discomfort if $\Delta p_{ear}>p_c$. The energy estimation presumes that the earplug is a perfect flow rate source, which is a theoretical hypothesis leading to $0.9$\,mJ available from one mastication. Based on 2200 mastication cycles per day, according to \cite{Goll2011}, there would be $2$\,mJ available per day, per ear. This amount of energy could theoretically enhance supply $23.5$\,\% of the $17$\,J of needed for the state of the art of cochlear implants \cite{Kulah2022}. Further research in both domains related to hearing aids and energy harvesting could lead to fully autonomous devices in the future.
\begin{align}
	\text{max}(E_{hyd}) = p_c \cdot (\Delta V_{ear})_{max} = 0.9\text{\,mJ}
	\label{eq:max_hydraulic_energy}
\end{align}

    %///////////////////////////////////////////// 
	\subsection{The harvesting strategy}	
	\label{The harvesting strategy}
    %/////////////////////////////////////////////
Electromagnetic transducers are usually used to harvest energy from kinetic (vibration) energy sources. In fact, the power density of electromagnetic transducers decreases as much as the effective volume of transducing material and the frequency of the energy sources decreases \cite{Priya2017,Kulah2008}. Also, the fabrication and the integration of the coil and the moving magnet can come out to be difficult as the gap between the two components has to be carefully managed to prevent non-linear unpredicted responses \cite{Caruntu2001}. Thus, piezoelectric ceramics appear like good candidates, but they cannot be easily used because of their high stiffness and their high resonance frequencies (>$1.57$\,Hz). Therefore, the harvesting system could benefit from a frequency-up conversion stage \cite{Ashraf2011,Peng2021} to increase the frequency of the energy source and so the energy conversion efficiency of piezoelectric ceramic transducers. A majority of the frequency-up systems with piezoelectric transducers use mechanical stops with piezoelectric ceramic cantilevers to reach high frequencies with a good mechanical coupling \cite{Edwards2013,Gu2011,Lee2007}. Stoppers, however, dissipate energy. Alternative solutions use mechanically bistable resonators (BR) to perform frequency-up \cite{Vocca2012}. Recent works led to the development and optimization of  bistable structures integrating stacked piezoelectric ceramics inside flextensional elastic structures \cite{Huguet2017}. They show enhanced performances due to high electromechanical coupling.

We can estimate the best choice maximizing the energy extraction from the hydraulic source. To be harvested, the energy has to be stored in a mechanical potential reservoir at first, to be then converted in electricity by the transducer. The hydraulic-mechanic interface can be a hydraulic cylinder (HC). It could store then the energy in the frequency-up converter by actuating it at the frequency of human mastication. Figure \ref{fig:hydraulic_mechanic_interface} schematizes the mechanical interaction between the HC and the resonator assimilated to a spring of stiffness $K_s$. During the low frequency actuation, the forces applied on each side of the HC piston are equal. Equation \ref{eq:hydraulic-mechanic_interface} models the hydraulic-mechanic interface with $x_{m}$ and $ S_{hc}$ being respectively the dynamic mass position on the $\vec{x}$ axis and the HC internal section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
	\Delta p_{ear}\ S_{hc} = K_{s}\ x_{m}
	\label{eq:hydraulic-mechanic_interface}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={19.7cm 0cm 0cm 13cm},clip, width=0.7\linewidth]{figures/hydraulic_mechanic_interface.pdf}
	\caption{Hydraulic source exploitation depending on the mechanical load}
	\label{fig:hydraulic_mechanic_interface}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Therefore, the potential energy stored in a linear resonator from the ideal hydraulic source previously described can be expressed as follows :
\begin{equation}
	Ep_{lin} = \dfrac{K_s\ x_{m}^2}{2} = \dfrac{\Delta p_{ear} \Delta V_{ear}}{2}
	\label{eq:best_linear_soution}
\end{equation}
Thus, the linear solution can theoretically exploit/store a maximum of $50\,\%$ from the maximum available hydraulic energy ($\text{max}(E_{hyd})$).

Yet, BRs usually admit non-linear elastic behaviour. For example, a BR governed by a Duffing equation is inherently capable of exploiting $65$\,\% from the same energy source, which is $15$\,\% better than the best theoretical linear option. Section \ref{sec:SYSTEM MODELING} discusses the details of the modelling and design of the harvester that implements such a non-linear BR as it appeared to be a better solution than the linear approach.

% Figure \ref{fig:OB_vs_LINEAR} shows in blue the hydraulic behaviour of the ideal flow rate source previously introduced with the upper limit values of pressure and volume variations.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}[!htbp]
% 	\centering
% 	\captionsetup{justification=centering}
% 	\includegraphics[trim={22.5cm 0cm 0cm 8.8cm},clip, width=0.35\textwidth]{figures/OB_vs_LINEAR.pdf}
% 	\caption{Hydraulic source exploitation depending on the mechanical load}
% 	\label{fig:OB_vs_LINEAR}
% \end{figure}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{HARVESTER PRESENTATION AND OPERATION \mbox{PRINCIPLE}}
\label{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
